Quantifying and Modeling Global Water Accessibility including Water Typology

### 2021-01-21 docs/manuscript/manuscript.tex
Updates from Overleaf
Edit of labels for figures.
Insertion of cluster-analysis figures.

Table 3 is a rough summary of why Ward.D2 is the best clustering methodology if we go by four cuts. For the KL cluster algorithm, we determined that the best Value index is 2.4982 which is the lowest for the index, and the KL algorithm determines that the lowest value index results from the best cut of four. Also, the Ratkowsky cluster theorem determined that Ward.D2 was the best method for four cuts because the value index was the lowest matching Ward.D at .2981. However, Ward.D is the incomplete version of Ward.D2 where they forgot to square one of the intermediate value in the theorem. Therefore, Ward.D2 is the best clustering method for Ward.D2. Similarly, for the Marriot clustering theorem, Ward.D2 is the best once again due to the Ward.D having the greatest value index which determines that it is the best clustering method. Similar to an explanation from Ratkowsky, Ward.D2 is chosen to be the best method for 4 cuts.

@@

Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables. The majority of the data is from 2010 - 2020 (Figure 1). Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys. Most of the variables are related to water accessibility by their locations, time, and water resources. This also allows us to find water accessibility predictor variables and new clustering graphs.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{correlation-basic}
  \caption{Correlation Graph of Numerical Variables}
  \label{fig:year}
\end{figure}
As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p-value, and to see a relationship between different categorical variables is shown in Figure 3, which again indicates a greater detailed comparison between the variables. Households with an improved water source on the premises (iwsp) have a strong relationship with households with basic water service (basw) of .84 and a p-value of less than 0.001. Also, we learned and discovered that Households using an unimproved water source (uiws) have a positive relationship with Households using unprotected well water (uwel), Households using an unprotected spring (uspr), and Households using surface water (surw) with a p-value of less than 0.001 as well. These relationships justify the necessity to research further if the particular water accessibility variables can increase the accessibility of particular water sources to reach universal access to water by 2030, as proposed in the SDGs.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{adv-cor-plot}
  \caption{Advanced Correlation Graph of Numerical Variables}
  \label{fig:year}
\end{figure}
Additionally, we wanted to explore if the modes of transportation variables have any relationship with potential water variables. Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which are shown in Figure 4,5,6. There is a significant positive relationship between households possessing a private car (pcar) and households with water on the premises (watp) as shown in Figure 4. There is also is a significant negative relationship between households possessing a private car (pcar) with households with water 30 minutes or less away round trip (tles) and households with water more than 30 minutes away round trip (tmor) with a p-value of less than 0.001. This information is essential to investigate further why this is the global trend.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-time}
  \caption{Advanced Correlation Graph of Modes of Transportation and Time}
  \label{fig:year}
\end{figure}
Figure 5 shows that Households possessing a motorcycle (mcyc) has a positive correlation with Households using a tube well/borehole (bore) and Households using rainwater (rain) with a p-value of less than 0.01. Similar to the previous analysis, there is a relationship between a mode of transportation to retrieve potential water sources. Figure 6 also shows that households possessing an animal-drawn cart (acar) have a positive and strong relationship with households using unprotected well water (uwel) with a p-value of less than 0.001. The results demonstrate that different modes of transportation factor into the accessibility of retrieving water sources on a global scale, therefore, there should be consideration of these variables to increase access to improved water sources. 
\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-water-1}
  \caption{Advanced Correlation Graph of Modes of Transportation and Water Resources}
  \label{fig:year}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-water-2}
  \caption{Advanced Correlation Graph of Modes of Transportation and Additional Water Resources}
  \label{fig:year}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{spider-plot}
  \caption{Spider Plot of Water Accessibility Variables by Typology}
  \label{fig:spider}
\end{figure}

>>>>

Table 3 is a rough summary of why Ward.D2 is the best clustering methodology if we go by four cuts. For the KL cluster algorithm, we determined that the best Value index is 2.4982 which is the lowest for the index, and the KL algorithm determines that the lowest value index results from the best cut of four. Also, the Ratkowsky cluster theorem determined that Ward.D2 was the best method for four cuts because the value index was the lowest matching Ward.D at .2981. However, Ward.D is the incomplete version of Ward.D2 where they forgot to square one of the intermediate value in the theorem. Therefore, Ward.D2 is the best clustering method for Ward.D2. Similarly, for the Marriot clustering theorem, Ward.D2 is the best once again due to the Ward.D having the greatest value index which determines that it is the best clustering method. Similar to an explanation from Ratkowsky, Ward.D2 is chosen to be the best method for 4 cuts. Figure 7 shows the visual representation of the explanation by the table 3.

@@

Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables. The majority of the data is from 2010 - 2020 as shown in Figure 1. Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys. Most of the variables are related to water accessibility by their locations, time, and water resources. This also allows us to find water accessibility predictor variables and new clustering graphs.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{correlation-basic}
  \caption{Correlation Graph of Numerical Variables}
  \label{fig:cor-basic}
\end{figure}
As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p-value, and to see a relationship between different categorical variables is shown in Figure 3, which again indicates a greater detailed comparison between the variables. Households with an improved water source on the premises (iwsp) have a strong relationship with households with basic water service (basw) of .84 and a p-value of less than 0.001. Also, we learned and discovered that Households using an unimproved water source (uiws) have a positive relationship with Households using unprotected well water (uwel), Households using an unprotected spring (uspr), and Households using surface water (surw) with a p-value of less than 0.001 as well. These relationships justify the necessity to research further if the particular water accessibility variables can increase the accessibility of particular water sources to reach universal access to water by 2030, as proposed in the SDGs.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{adv-cor-plot}
  \caption{Advanced Correlation Graph of Numerical Variables}
  \label{fig:cor-plot}
\end{figure}
Additionally, we wanted to explore if the modes of transportation variables have any relationship with potential water variables. Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which are shown in Figure 4,5,6. There is a significant positive relationship between households possessing a private car (pcar) and households with water on the premises (watp) as shown in Figure 4. There is also is a significant negative relationship between households possessing a private car (pcar) with households with water 30 minutes or less away round trip (tles) and households with water more than 30 minutes away round trip (tmor) with a p-value of less than 0.001. This information is essential to investigate further why this is the global trend.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-time}
  \caption{Advanced Correlation Graph of Modes of Transportation and Time}
  \label{fig:transporation-time}
\end{figure}
Figure 5 shows that Households possessing a motorcycle (mcyc) has a positive correlation with Households using a tube well/borehole (bore) and Households using rainwater (rain) with a p-value of less than 0.01. Similar to the previous analysis, there is a relationship between a mode of transportation to retrieve potential water sources. Figure 6 also shows that households possessing an animal-drawn cart (acar) have a positive and strong relationship with households using unprotected well water (uwel) with a p-value of less than 0.001. The results demonstrate that different modes of transportation factor into the accessibility of retrieving water sources on a global scale, therefore, there should be consideration of these variables to increase access to improved water sources. 
\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-water-1}
  \caption{Advanced Correlation Graph of Modes of Transportation and Water Resources}
  \label{fig:transporation-water}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-water-2}
  \caption{Advanced Correlation Graph of Modes of Transportation and Additional Water Resources}
  \label{fig:transporation-water}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{optimal-cuts}
  \caption{Optimal Number of Clusters for Ward.D2}
  \label{fig:optimal}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{centroid}
  \caption{Centroid of the Four Cluster Groupings}
  \label{fig:centroid}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{typology}
  \caption{Cluster Typology Average of Water Accessibility Variables }
  \label{fig:typology}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{dendrogram}
  \caption{Cluster Dendrogram with Four Cuts}
  \label{fig:dendrogram}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{spider-plot}
  \caption{Spider Plot of Water Accessibility Variables by Typology}
  \label{fig:spider}
\end{figure}
